---
title: "Comparing Dissmilarity Indices using Current and Historical Metro Definitions"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: "header.html"
      after_body: "./footer.html"
---

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #4b2e83;
}
</style>

The goal of this research is to assess the sensitivity of measuring historical metropolitan segregation using present-day Core-Based Statistical Areas (CBSAs) in contrast to contemporaneous boundaries (for example, 1970 Standard Metropolitan Statistical Areas). The Census Bureau has changed the extent of what constitutes a metropolitan area in a given year over the past 40 years, and these boundary revisions complicate measuring metropolitan residential segregation over long periods of time (for example, 1970 to 2010). 

<br>

This problem is conceptually similar to when census tract definitions (a common areal summary for neighborhoods) change between data collections. In this case, differences between the two measurements of total population reflect both _population change_ (which is likely a quantity of interest) but also _measurement error_ created by summarizing overlapping but not identical areas for which persons can exist. The typical approach is to normalize boundaries such that the same set of space is aggregated across different points in time. The units themselves may still be somewhat arbitrary in how they divide up space so long as they are comparable.

<br>
<hr>
<br>

## Preamble

We'll start by loading the `tidyverse` package that covers `dplyr` (data manipulation), `ggplot` (data visualization), `stringr` (string manipulation) and `readr` (data import). `skimr` provides pipable functions for summary statistics that will be useful for analyzing agreement between measurement strategies.

```{r library, warning=FALSE, message=FALSE}
#libraries
library(tidyverse)
library(skimr)
```

Loading the data for this analysis requires two different steps. The first step is to load the NCDB into the R environment (NB: see [here](http://www.geolytics.com/USCensus,Neighborhood-Change-Database-1970-2000,Products.asp) for more information about obtaining a copy from Geolytics). Since this file is a .RData object, we use `load("path_to_filename.RData").` The second step is to load in comma-separated value files for the Office of Management and Budget Metropolitan Area delineations corresponding to 1970, 1980, 1990 and 2000. 

```{r data, warning=FALSE, message=FALSE}
#load in NCDB to extract CBSAs with (NB: RData untracked on github)
load(file = "./ncdb_workspace.RData")

#load in OMB definition file to use as crosswalk
msa1970 <- read_csv("input/71mfips.csv") #delimited county-MSA data for
msa1980 <- read_csv("input/80mfips.csv") #each census year
msa1990 <- read_csv("input/90mfips.csv")
msa2000 <- read_csv("input/99mfips.csv")
```

Lastly, we save a function in the form of an operator like `>` and `==` with `%!in%`, which returns `TRUE` when LHS values being evaluated are not found in a value or set of values on the RHS. This is useful later for filtering places that are not in New England given the data differences for MSAs in these states.

```{r NOT IN}
#not in operator
'%!in%' <- function(x,y)!('%in%'(x,y))

#examples
primary <- c("Blue", "Red", "Yellow")

"Blue" %in% primary

"Green" %in% primary

"Green" %!in% primary
```

<br>
<hr>
<br>

## Transform OMB and NCDB tables

The next step is to transform the starting tables to the structure needed. For each OMB table, we would like a row to uniquely identify counties and to denote what metropolitan area code was tied to the county at a specific point in time. Unfortunately, New England states can have multiple MSAs to a county, so because of this situation we level the data to capture all unique county subdivisions. Outside of New England there can be multiple county subdivisions to a county, but definitions of MSAs do not vary within a county (i.e. a county has at most 1 valid MSA code).

The process for a given period follows these steps:

1. remove completely blank rows (denoted by not having a state value)
2. name the COUSUB variable (first valid line of each table has no value since not in New England)
3. pad zeroes on the ID codes for comparability
4. remove county subdivisions that did not have a MSA code for the specified year
5. group table by unique STATE X COUNTY X COUSUB combinations. In NE this varies by STATE X COUNTY X COUSUB, outside NE the table varies by STATE X COUNTY only
6. filter rows where: 
    + multiple SMSAXX codes exist and there is no town denoted (these are heading rows for each MSA in New England) OR
    + there's only one MSA value among the group (this is all of the counties outside of New England)
7. return the unique value found for each STATE X COUNTY X COUSUB combination (length 1 in each combination)

```{r Transform data 1970/1980, warning=FALSE}
#1970 SMSA definitions
msa1970 <- msa1970 %>% #wipe out the cases for the MSA alone --- they are coded
  filter(!is.na(STATE)) %>% #as missing STATE IDs
  mutate(COUSUB = X6) %>% #establish colname for COUSUB field
  select(STATE, COUNTY, COUSUB, SMSA70) %>% #remove other variables from data frame
  mutate(STATE = str_pad(STATE, 2, side = "left", pad = "0"), #pad ID strings with 0s as needed
         COUNTY = str_pad(COUNTY, 3, side = "left", pad = "0"),
         COUSUB = str_pad(COUSUB, 5, side = "left", pad = "0"),
         FIPS = paste0(STATE, COUNTY)) %>%
  filter(!is.na(SMSA70)) %>% #remove rows with no valid MSA code
  group_by(STATE, COUNTY, FIPS, COUSUB) %>% #group by unique State X Cty/Cty X Town combinations
  filter((length(unique(SMSA70)) != 1 & !is.na(COUSUB)) | #if two or more SMSA codes for the combination, keep those that correspond to a COUSUB in NE
           length(unique(SMSA70)) == 1) %>% #and keep all rows where there's only one SMSA code for the combination (i.e. one county, one SMSA)
  summarize(MSA70 = unique(SMSA70)) #for each combination, the summary value is the unique (i.e. only) code present

#1980 SMSA definitions
msa1980 <- msa1980 %>%
  filter(!is.na(STATE)) %>%
  mutate(COUSUB = X8) %>%
  select(STATE, COUNTY, COUSUB, SMSA80) %>%
  mutate(STATE = str_pad(STATE, 2, side = "left", pad = "0"),
         COUNTY = str_pad(COUNTY, 3, side = "left", pad = "0"),
         COUSUB = str_pad(COUSUB, 5, side = "left", pad = "0"),
         FIPS = paste0(STATE, COUNTY)) %>%
  filter(!is.na(SMSA80)) %>%
  group_by(STATE, COUNTY, FIPS, COUSUB) %>%
  filter((length(unique(SMSA80)) != 1 & !is.na(COUSUB)) |
           length(unique(SMSA80)) == 1) %>%
  summarize(MSA80 = unique(SMSA80))
```

For the 1990 and 2000 OMB delineations the process needs to deal with OMB's creation of _consolidated metropolitan statistical areas (CMSA)_. Large metropolitan areas were assigned a single CMSA code covering multiple distinct housing and employment markets, with these places' metropolitan units most comparable to the 1970 and 1980 _standard metropolitan statistical areas (SMSA)_ living on as _primary metropolitan statistical areas (PMSA)_ that exist within CMSAs. Thus, for the following code we process the data more or less the same up to the point of using `summarize()` to assign the unique MSA code observed for the STATE X COUNTY X COUSUB combination. 

For the 1990 and 2000 data aggregation with `summarize()`, we now need to conditionally assign either the PMSA code or the CMSA code. The column of the OMB file for CMSA has MSA values for all counties and county subdivisions that do not have a higher-oreder CMSA. When a higher-order CMSA exists, a separate field indicates the PMSA (otherwise missing as `NA`). `ifelse()` has three arguments---the first is the expression to test, the second is the value to assign to cases where the expression was `TRUE`, the last is the value to assign where the expression was `FALSE`. Accordingly, `ifelse(is.na(unique(PMSA90)), unique(CMSA90), unique(PMSA90)))` assigns the unique CMSA code in 1990 to cases where the value for PMSA in 1990 was missing, and otherwise uses the PMSA code if not missing.

```{r Transform data 1990/2000, warning=FALSE}
#1990 PMSA definitions
msa1990 <- msa1990 %>%
  filter(!is.na(STATE)) %>%
  mutate(COUSUB = X12) %>%
  select(STATE, COUNTY, COUSUB, PMSA90, CMSA90) %>%
  mutate(STATE = str_pad(STATE, 2, side = "left", pad = "0"),
         COUNTY = str_pad(COUNTY, 3, side = "left", pad = "0"),
         COUSUB = str_pad(COUSUB, 5, side = "left", pad = "0"),
         FIPS = paste0(STATE, COUNTY)) %>%
  filter(!is.na(CMSA90)) %>%
  group_by(STATE, COUNTY, FIPS, COUSUB) %>%
  filter((length(unique(PMSA90)) != 1 & !is.na(COUSUB)) |
           length(unique(PMSA90)) == 1) %>%
  summarize(MSA90 = ifelse(is.na(unique(PMSA90)), unique(CMSA90), unique(PMSA90))) 

#2000 PMSA definitions
msa2000 <- msa2000 %>%
  filter(!is.na(STATE)) %>%
  mutate(COUSUB = X12) %>%
  select(STATE, COUNTY, COUSUB, PMSA99, CMSA99) %>%
  mutate(STATE = str_pad(STATE, 2, side = "left", pad = "0"),
         COUNTY = str_pad(COUNTY, 3, side = "left", pad = "0"),
         COUSUB = str_pad(COUSUB, 5, side = "left", pad = "0"),
         FIPS = paste0(STATE, COUNTY)) %>%
  filter(!is.na(CMSA99)) %>%
  group_by(STATE, COUNTY, FIPS, COUSUB) %>%
  filter((length(unique(PMSA99)) != 1 & !is.na(COUSUB)) |
           length(unique(PMSA99)) == 1) %>%
  summarize(MSA00 = ifelse(is.na(unique(PMSA99)), unique(CMSA99), unique(PMSA99)))
```

After munging the OMB delineation files, we do some light cleaning to the vanilla NCDB object (i.e. the NCDB as it comes from Geolytics) to remove tracts outside of CBSAs and to pad FIPS IDs to match the same structure as in other tables. This census tract unit table will be used to identify what MSA a given tract was in at a specific point in time based on STATE X COUNTY or STATE X COUNTY X COUSUB combinations (if New England).

```{r Crosswalk}
#Crosswalk table of tracts with accompanying higher level IDs
ncdb_cw <- ncdb_vanilla %>%
  filter(CBSA != "") %>%
  mutate(STATE = str_pad(STATE, 2, side = "left", pad = "0"),
         COUNTY = str_pad(COUNTY, 3, side = "left", pad = "0"),
         COUSUB = str_pad(COUSUB, 5, side = "left", pad = "0"),
         FIPS = paste0(STATE, COUNTY))
```

<br>
<hr>
<br>

## Compile the crosswalk table

There are two sets of joins here. The first is for all tracts outside of New England and results in `ncdb_nonNE`, the second is for all New England tracts and creates `ncdb_NE`. Note how the filter function that is piped onto each object used in `left_join` handles this split by using `%in%` or `%!in%` matches made to the set of New England states with MSAs at the point in time. To ensure that the joins outside of New England are made according to STATE X COUNTY matches, we compute the join only after removing the COUSUB field from the table of segregation measures for each decade. The tracts from the NCDB may only match one set of values in the segregation table, so the resulting table of `bind_rows(ncdb_nonNE, ncdb_NE)` has the same number of observations as the `ncdb_cw` object we start with.

```{r Compile transformed objects, warning=FALSE}
start <- nrow(ncdb_cw)

#join MSA vals to NCDB cases outside of New England by COUNTY X STATE
ncdb_nonNE <- left_join(ncdb_cw %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")),
                        msa1970 %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")) %>% 
                          select(-COUSUB))

ncdb_nonNE <- left_join(ncdb_nonNE %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")),
                        msa1980 %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")) %>% 
                          select(-COUSUB)) 

ncdb_nonNE <- left_join(ncdb_nonNE %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")),
                        msa1990 %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")) %>% 
                          select(-COUSUB)) 

ncdb_nonNE <- left_join(ncdb_nonNE %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")),
                        msa2000 %>% filter(STATE %!in% c("09", "23", "25", "33", "44", "50")) %>% 
                          select(-COUSUB)) 

#join MSA vals to NCDB cases inside New England by COUNTY X STATE X COUSUB
ncdb_NE <- left_join(ncdb_cw %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")),
                     msa1970 %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")))

ncdb_NE <- left_join(ncdb_NE %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")),
                     msa1980 %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")))

ncdb_NE <- left_join(ncdb_NE %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")),
                     msa1990 %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")))

ncdb_NE <- left_join(ncdb_NE %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")),
                     msa2000 %>% filter(STATE %in% c("09", "23", "25", "33", "44", "50")))

#join NE and non-NE tables together
ncdb_cw <- bind_rows(ncdb_nonNE, ncdb_NE) %>%
  arrange(STATE, COUNTY)

end <- nrow(ncdb_cw)

start == end
```

<br>
<hr>
<br>

## Compute Dissimilarity Indices

At this point we can almost compute the segregation measures. However, since we want non-Hispanic black and non-Hispanic white counts as the two population groups we need to pull our work from the `ncdb_cw` object created from our vanilla NCDB table. We do this by extracting a table of tract, state, county, town and time-varying MSA IDs for each tract in the NCDB. This lets us merge time-varying MSA geography onto our interpolated NCDB where we also have non-Hispanic black and non-Hispanic white counts back till 1970 (obtained via allocation using Hispanic person count). After making the tract ID column name consistent across tables, we can append the IDs from `ncdb_cw` onto `ncdb_interp` and save a new table `ncdb` which we'll use from here out.

```{r Transfer NCDB IDs, warning=FALSE}
ncdb_cw <- ncdb_cw %>%
  ungroup() %>%
  distinct(GEO2010, STATE, COUNTY, COUSUB, MSA70, MSA80, MSA90, MSA00, CBSA) #really just need tract ID + MSA IDs

ncdb_interp$GEO2010 <- ncdb_interp$geo2010

ncdb <- inner_join(ncdb_cw, ncdb_interp)
```

Computing the dissimilarity index takes four piped functions. Starting with the tract-level `ncdb`:

1. `filter(!is.na(MSAXX))` - remove tracts that don't exist within the selected boundary
2. `group_by(MSAXX)` - group tracts by the selected boundary
3. `summarize(MSAXX_D_19XX = * sum(abs(tnhbt1970/sum(tnhbt1970) - tnhwt1970/sum(tnhwt1970))))` - summarize each MSA's segregation as $$\frac{1}{2} \times \sum_{i=1}^{N}{|\frac{\textrm{black}_{i}}{\textrm{black}_{metro}}-\frac{\textrm{white}_{i}}{\textrm{white}_{metro}}}|$$ where i are tracts and there are N per MSA grouping.
4. `mutate_at()`- this is mutating the column made via summarize to recode as `NA` in the places where division by zero took place (creates `NaN` instead).

For the historical delineations we only compute the contemporaneous measure. For the 2010 CBSA definition, we compute measures of dissmilarity for 1970-2000.

```{r Compute Dissimilarity, warning=FALSE}
seg_1970msa <- ncdb %>%
  filter(!is.na(MSA70)) %>%
  group_by(MSA70) %>%
  summarize(MSA70_D_1970 = (.5) * sum(abs(tnhbt1970/sum(tnhbt1970) - tnhwt1970/sum(tnhwt1970)))) %>%
  mutate_at(.vars = vars(starts_with("MSA70_D")),
            .funs = function(x){ifelse(is.nan(x), NA, x)})

seg_1980msa <- ncdb %>%
  filter(!is.na(MSA80)) %>%
  group_by(MSA80) %>%
  summarize(MSA80_D_1980 = (.5) * sum(abs(tnhbt1980/sum(tnhbt1980) - tnhwt1980/sum(tnhwt1980)))) %>%
  mutate_at(.vars = vars(starts_with("MSA80_D")),
            .funs = function(x){ifelse(is.nan(x), NA, x)})

seg_1990msa <- ncdb %>%
  filter(!is.na(MSA90)) %>%
  group_by(MSA90) %>%
  summarize(MSA90_D_1990 = (.5) * sum(abs(tnhbt1990/sum(tnhbt1990) - tnhwt1990/sum(tnhwt1990)))) %>%
  mutate_at(.vars = vars(starts_with("MSA90_D")),
            .funs = function(x){ifelse(is.nan(x), NA, x)})

seg_2000msa <- ncdb %>%
  filter(!is.na(MSA00)) %>%
  group_by(MSA00) %>%
  summarize(MSA00_D_2000 = (.5) * sum(abs(tnhbt2000/sum(tnhbt2000) - tnhwt2000/sum(tnhwt2000)))) %>%
  mutate_at(.vars = vars(starts_with("MSA00_D")),
            .funs = function(x){ifelse(is.nan(x), NA, x)})

seg_2010cbsa <- ncdb %>%
  filter(!is.na(CBSA)) %>%
  group_by(CBSA) %>%
  summarize(CBSA_D_1970 = (.5) * sum(abs(tnhbt1970/sum(tnhbt1970) - tnhwt1970/sum(tnhwt1970))),
            CBSA_D_1980 = (.5) * sum(abs(tnhbt1980/sum(tnhbt1980) - tnhwt1980/sum(tnhwt1980))),
            CBSA_D_1990 = (.5) * sum(abs(tnhbt1990/sum(tnhbt1990) - tnhwt1990/sum(tnhwt1990))),
            CBSA_D_2000 = (.5) * sum(abs(tnhbt2000/sum(tnhbt2000) - tnhwt2000/sum(tnhwt2000)))) %>%
  mutate_at(.vars = vars(starts_with("CBSA_D")),
            .funs = function(x){ifelse(is.nan(x), NA, x)})

#NB: NAs in the 2010 CBSA table come from two sources
#1. tracts with 0 population at t -> CBSAs with 0 population at t
#2. NAs for tract population at t -> NA for dissimilarity at t
```

After computing individual tables with MSA codes and dissimilarity we join the historical-delineated and CBSA-delineated measures to a county subdivision table so that each row uniquely identifies a areal summary. Within New England there is variation within counties but elsewhere the data only vary by unique counties and can be safely collapsed. Running an analysis on county level data (i.e. where a given STATE X COUNTY combination instead appears as many times as it belongs to unique MSAs in New England, but there is no COUSUB unit to the data) produces substantively similar results.

```{r Compile seg_D}
#compute crosswalk table of counties according to the 
seg_cw_NE <- ncdb_cw %>%
  mutate(NEWENG = STATE %in% c("09", "23", "25", "33", "44", "50")) %>%
  filter(NEWENG) %>%
  distinct(STATE, COUNTY, COUSUB, MSA70, MSA80, MSA90, MSA00, CBSA, NEWENG)

seg_cw_nonNE <- ncdb_cw %>%
  mutate(NEWENG = STATE %in% c("09", "23", "25", "33", "44", "50")) %>%
  filter(!NEWENG) %>%
  mutate(COUSUB = NA) %>% #remove variation in this to allow
  distinct(STATE, COUNTY, COUSUB, MSA70, MSA80, MSA90, MSA00, CBSA, NEWENG)

seg_cw <- bind_rows(seg_cw_NE, seg_cw_nonNE)

#save crosswalk file to storage
write_csv(seg_cw, "./output/county_msa_cbsa_crosswalk.csv")

#append segregation measures by CBSA values
seg_D <- left_join(seg_cw, seg_2010cbsa)  
seg_D <- left_join(seg_D, seg_2000msa)
seg_D <- left_join(seg_D, seg_1990msa)
seg_D <- left_join(seg_D, seg_1980msa)
seg_D <- left_join(seg_D, seg_1970msa)
```

<br>
<hr>
<br>

## Analyze measurement differences

```{r Analysis, warning=FALSE}
#number/prop of historical MSA counties in CBSA by decade
cbsa_sum <- seg_D %>%
  group_by(CBSA, NEWENG) %>%
  summarize(nCBSA = n(),
            nMSA70 = sum(!is.na(MSA70)),
            nMSA80 = sum(!is.na(MSA80)),
            nMSA90 = sum(!is.na(MSA90)),
            nMSA00 = sum(!is.na(MSA00))) %>%
  mutate(prop70 = nMSA70/nCBSA,
         prop80 = nMSA80/nCBSA,
         prop90 = nMSA90/nCBSA,
         prop00 = nMSA00/nCBSA,
         micro = (nMSA70 + nMSA80 + nMSA90 + nMSA00) == 0) #if county is not associated with metro area code through 70-00, must be micropolitan MSA cty added in 2010

#join summaries about each CBSA's counties back to seg_D by each county's CBSA code
seg_D <- left_join(seg_D, cbsa_sum)

#set filter for time (i.e. same MSAs across time or add in by decade)
seg_D <- seg_D %>%
  filter(!micro) #keep counties in CBSAs that had at least one county in the equivalent 1970 MSA

#spruce up indicator for county subdivisions in NE states
seg_D$NEWENG <- factor(seg_D$NEWENG)
levels(seg_D$NEWENG) <- c("Outside N.E.", "New England")
```

### Share of historical MSA counties in current CBSAs

#### 1970

```{r Tables Prop Cty 70, message=FALSE, results="asis"}
cbsa_sum %>%
  filter(!micro) %>%
  group_by(NEWENG) %>%
  select(NEWENG, prop70) %>%
  skim() %>%
  skimr::kable()
```

#### 1980

```{r Tables Prop Cty 80, message=FALSE, results="asis"}
cbsa_sum %>%
  filter(!micro) %>%
  group_by(NEWENG) %>%
  select(NEWENG, prop80) %>%
  skim() %>%
  skimr::kable()
```

#### 1990

```{r Tables Prop Cty 90, message=FALSE, results="asis"}
cbsa_sum %>%
  filter(!micro) %>%
  group_by(NEWENG) %>%
  select(NEWENG, prop90) %>%
  skim() %>%
  skimr::kable()
```

#### 2000

```{r Tables Prop Cty 00, message=FALSE, results="asis"}
cbsa_sum %>%
  filter(!micro) %>%
  group_by(NEWENG) %>%
  select(NEWENG, prop00) %>%
  skim() %>%
  skimr::kable()
```

### Correlation of D using historical and CBSA delineations

```{r Tables Cor, message=FALSE, results="asis"}
knitr::kable(seg_D %>% 
  group_by(NEWENG) %>% 
  summarize(cor70 = cor(MSA70_D_1970, CBSA_D_1970, use = "complete.obs"),
            cor80 = cor(MSA80_D_1980, CBSA_D_1980, use = "complete.obs"),
            cor90 = cor(MSA90_D_1990, CBSA_D_1990, use = "complete.obs"),
            cor00 = cor(MSA00_D_2000, CBSA_D_2000, use = "complete.obs")))
```

### Summary statistics of D distributions

#### 1970

```{r Tables Summary 70, message=FALSE, results="asis"}
seg_D %>% 
  group_by(NEWENG) %>%
  select(MSA70_D_1970, CBSA_D_1970) %>% 
  skim() %>%
  skimr::kable()
```

#### 1980

```{r Tables Summary 80, message=FALSE, results="asis"}
seg_D %>% 
  group_by(NEWENG) %>%
  select(MSA80_D_1980, CBSA_D_1980) %>% 
  skim() %>%
  skimr::kable()
```

#### 1990

```{r Tables Summary 90, message=FALSE, results="asis"}
seg_D %>% 
  group_by(NEWENG) %>%
  select(MSA90_D_1990, CBSA_D_1990) %>% 
  skim() %>%
  skimr::kable()
```

#### 2000

```{r Tables Summary 00, message=FALSE, results="asis"}
seg_D %>% 
  group_by(NEWENG) %>%
  select(MSA00_D_2000, CBSA_D_2000) %>% 
  skim() %>%
  skimr::kable()
```

### Counties with exact same and within value for MSA and CBSA measures

#### 1970

```{r Tables Same/Within 70, message=FALSE, results="asis"}
knitr::kable(seg_D %>% 
  filter(!is.na(MSA70_D_1970), !is.na(CBSA_D_1970)) %>%
  mutate(same = MSA70_D_1970 == CBSA_D_1970,
         within_01 = abs(MSA70_D_1970 - CBSA_D_1970) <= .01,
         within_05 = abs(MSA70_D_1970 - CBSA_D_1970) <= .05) %>%
  group_by(NEWENG) %>%
  summarize(nSame = sum(same),
            nWithin_01 = sum(within_01),
            nWithin_05 = sum(within_05),
            nTot = n(),
            propSame = nSame/nTot,
            propWithin_01 = nWithin_01/nTot,
            propWithin_05 = nWithin_05/nTot)) 
```

#### 1980

```{r Tables Same/Within 80, message=FALSE, results="asis"}
knitr::kable(seg_D %>% 
  filter(!is.na(MSA80_D_1980), !is.na(CBSA_D_1980)) %>%
  mutate(same = MSA80_D_1980 == CBSA_D_1980,
         within_01 = abs(MSA80_D_1980 - CBSA_D_1980) <= .01,
         within_05 = abs(MSA80_D_1980 - CBSA_D_1980) <= .05) %>%
  group_by(NEWENG) %>%
  summarize(nSame = sum(same),
            nWithin_01 = sum(within_01),
            nWithin_05 = sum(within_05),
            nTot = n(),
            propSame = nSame/nTot,
            propWithin_01 = nWithin_01/nTot,
            propWithin_05 = nWithin_05/nTot)) 
```

#### 1990

```{r Tables Same/Within 90, message=FALSE, results="asis"}
knitr::kable(seg_D %>% 
  filter(!is.na(MSA90_D_1990), !is.na(CBSA_D_1990)) %>%
  mutate(same = MSA90_D_1990 == CBSA_D_1990,
         within_01 = abs(MSA90_D_1990 - CBSA_D_1990) <= .01,
         within_05 = abs(MSA90_D_1990 - CBSA_D_1990) <= .05) %>%
  group_by(NEWENG) %>%
  summarize(nSame = sum(same),
            nWithin_01 = sum(within_01),
            nWithin_05 = sum(within_05),
            nTot = n(),
            propSame = nSame/nTot,
            propWithin_01 = nWithin_01/nTot,
            propWithin_05 = nWithin_05/nTot)) 
```

#### 2000

```{r Tables Same/Within 00, message=FALSE, results="asis"}
knitr::kable(seg_D %>% 
  filter(!is.na(MSA00_D_2000), !is.na(CBSA_D_2000)) %>%
  mutate(same = MSA00_D_2000 == CBSA_D_2000,
         within_01 = abs(MSA00_D_2000 - CBSA_D_2000) <= .01,
         within_05 = abs(MSA00_D_2000 - CBSA_D_2000) <= .05) %>%
  group_by(NEWENG) %>%
  summarize(nSame = sum(same),
            nWithin_01 = sum(within_01),
            nWithin_05 = sum(within_05),
            nTot = n(),
            propSame = nSame/nTot,
            propWithin_01 = nWithin_01/nTot,
            propWithin_05 = nWithin_05/nTot)) 
```

<br>
<hr>
<br>

## Visualize measurement differences

#### 1970

```{r Visualization 1970, echo=FALSE, warning=FALSE, fig.width=6, fig.height=3, dpi=300}
#AvP-esque scatterplots
ggplot(seg_D, aes(x = MSA70_D_1970, y = CBSA_D_1970)) +
  facet_grid(~ NEWENG) +
  geom_hex() +
  geom_abline(slope = 1, intercept = 0) +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_viridis_c(option = "A") +
  labs(x = "1970 SMSA D for 1970 Population",
       y = "2010 CBSA D for 1970 Population")
```

<br>
<hr>
<br>

#### 1980

```{r Visualization 1980, echo=FALSE, warning=FALSE, fig.width=6, fig.height=3, dpi=300}
ggplot(seg_D, aes(x = MSA80_D_1980, y = CBSA_D_1980)) +
  facet_grid(~ NEWENG) +
  geom_hex() +
  geom_abline(slope = 1, intercept = 0) +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_viridis_c(option = "A") +
  labs(x = "1980 SMSA D for 1980 Population",
       y = "2010 CBSA D for 1980 Population")
```

<br>
<hr>
<br>

#### 1990

```{r Visualization 1990, echo=FALSE, warning=FALSE, fig.width=6, fig.height=3, dpi=300}
ggplot(seg_D, aes(x = MSA90_D_1990, y = CBSA_D_1990)) +
  facet_grid(~ NEWENG) +
  geom_hex() +
  geom_abline(slope = 1, intercept = 0) +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_viridis_c(option = "A") +
  labs(x = "1990 PMSA D for 1990 Population",
       y = "2010 CBSA D for 1990 Population")
```

<br>
<hr>
<br>

#### 2000

```{r Visualization 2000, echo=FALSE, warning=FALSE, fig.width=6, fig.height=3, dpi=300}
ggplot(seg_D, aes(x = MSA00_D_2000, y = CBSA_D_2000)) +
  facet_grid(~ NEWENG) +
  geom_hex() +
  geom_abline(slope = 1, intercept = 0) +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_viridis_c(option = "A") +
  labs(x = "2000 PMSA D for 2000 Population",
       y = "2010 CBSA D for 2000 Population")
```

